\begin{abstract}
% The \sysname tool enables humans as well as human-level AI to turn their logical constraints into time and energy efficient code.
% 
% The abstract serves as an executive summary of your project.
% Your abstract should cover at least the following topics, 1-2 sentences for
% each: what area you are in, the problem you focus on, why existing work is
% insufficient, what the high-level intuition of your work is, maybe a neat
% design or implementation decision, and key results of your evaluation.

GraalVM Native Image compiles Java programs into native binaries that operate under the closed-world
assumption. This means that every reﬂectively-accessed element such as Class, Executable, or Field
must be provided as reachability metadata to the image-build process. In Native Image, the reachability metadata is either 1) provided via user-provided conﬁguration ﬁles in JSON format or 2) computed by partially evaluating the input program to pre-compute the reﬂectively-accessed elements.

Due to the long image-build times, creating the correct metadata can be a time-consuming process for
users: if the metadata for any element is missing, the entire image needs to be rebuild.
The objective of this project is to allow users to compute metadata with a quick turnaround by
introducing a new mode to the JDK, so that it operates under the closed-world assumption and behaves
exactly like Native Image.

We modify all reﬂective JDK features to operate under a closed-world assumption by checking the
reachability metadata to determine if the reﬂectively accessed element is in the closed world. To pre-compute the reﬂectively-accessed elements we implement a bytecode-to-bytecode partial evaluator that transforms classes before they are loaded and extracts reachability metadata from them. Finally, implementing the Java mode is a mean to drive Native Image specifications. It enables us to clearly state what the expected behaviour is for all reflective accesses, and know which unexpected behaviours emerge due to compiler optimisations.   
\end{abstract}

